---
title: 'Comparing to Analytic'
Section: 'Determining Distribution'
Author: "Adam Dede"
Course: "Data to Decisions: Master Statistics with R and tidyverse"
Course_URL: "https://www.udemy.com/course/statsr_ad" 
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

gitHubPath <- file.path(path.expand("~"), "GitHub", 
                        "MasterStatsUsingR")


library(tidyverse) # data manipulation and visualization
source(file.path(gitHubPath, 'courseTheme.R'))
library(pracma)

```

## human EEG data 

```{r EEG}

#Key variable name: pow_delta_O
df = read.csv(file.path(gitHubPath, 'data',
                          'restingEEG.csv'))

df$dummy = as.factor(1) 

df = df %>% filter(pow_delta_O < 250)

ggplot(df, aes(x = pow_delta_O, color = dummy, fill = dummy)) +
  geom_histogram() + myTheme + brightCol  + brightFil +
  theme(legend.position = 'none')


```

## human EEG description

```{r descriptives for EEG}

EEGsum <- df %>% summarize(setName = 'deltaPow', 
                           meanVal = mean(pow_delta_O), 
                           sdVal = sd(pow_delta_O), 
                           logMeanVal = mean(log(pow_delta_O)), 
                           logSdVal = sd(log(pow_delta_O)), 
                           medianVal = median(pow_delta_O))
                               



```

## human EEG against analytic distributions 

```{r analytic comparison for EEG}
#how many observations would we expect at different 
#EEG values? 
x <- df$pow_delta_O
analyticDists <- data.frame(
                            'xVal' = x)
                            
  



```

## make it into a function for ease of use

```{r functionitize}

plotAgainstDists <- function(df, ){

  x <- df[[]]

 

}


```

## my emails per day 

```{r email}

df = read.csv(file.path(gitHubPath, 'data',
                      'emailTimes.csv'))


df <- df %>% group_by(DayOfWeek, Month, DayOfMonth, Year) %>% 
  summarize(count = n(), .groups = 'drop')





```
## emails description

```{r descriptives for emails}

emailSum <- df %>% summarize(



```

## email frequency against analytic distributions 

```{r analytic comparison for email frequency}

plotAgainstDists(df, "count",  emailSum, 'email')


```

## sp 500 daily change 

```{r sp 500}

df = read.csv(file.path(gitHubPath, 'data', 
                        'SP500HistoricalData.csv'))
names(df)[6] <- 'Change'

df$Change = unlist(lapply(df$Change, function(x) 
                as.numeric(strsplit(x, '%')[[1]])))



```

## sp 500 description

```{r descriptives for sp500}

sp500Sum <- df %>% summarize(




```

## sp 500 changes against analytic distributions 

```{r analytic comparison for sp500}


plotAgainstDists(df, "Change", sp500Sum, 'sp500')


```

## human memory performance

```{r human memory}
df = read.csv(file.path(gitHubPath, 'data',
              'MemoryData.csv'))

df <- df %>% filter(TargALL_UVSD_d >=0)




```

## human memory description

```{r descriptives for human mem}

memSum <- df %>% summarize(



```

## human memory against analytic distributions 

```{r analytic comparison for human memory}

plotAgainstDists(df, "TargALL_UVSD_d",memSum, "mem data", 20)


```
