---
title: 'describing real data'
author: "Adam Dede"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

gitHubPath = 'C:\\Users\\Adam\\Documents\\GitHub\\MasterStatsUsingR\\'

library(tidyverse) # data manipulation and visualization
source(paste0(gitHubPath, 'courseTheme.R'))
library(pracma)

```

## human EEG data 

```{r EEG}
#load in the restingEEG dataset. There are a lot of variables in this dataset
#for this analysis let's just focus on the delta power in the occipital region 
#of the head. 
#Key variable name: pow_delta_O
df = read.csv(paste0(gitHubPath, '\\data\\restingEEG.csv'))

df$dummy = as.factor(1) 
df <- df %>% filter(pow_delta_O < 250, pow_delta_O > 0)

ggplot(df, aes(x = pow_delta_O, color = dummy, fill = dummy)) + 
  geom_histogram() + 
  myTheme + brightFil + brightCol + theme(legend.position = 'none')

```

## human EEG description

```{r descriptives for EEG}

EEGsum <- df %>% summarize(setName = 'deltaPow', 
                               meanVal = mean(pow_delta_O), 
                               medianVal = median(pow_delta_O),  
                               sdVal = sd(pow_delta_O), 
                               logMean = mean(log(pow_delta_O)),
                               logSD = sd(log(pow_delta_O)), 
                               offSet = 0) 



```

## human EEG against analytic distributions 

```{r analytic comparison for EEG}

x <- df$pow_delta_O
analyticDists <- data.frame(
                            'norm' = dnorm(x, 
                                           mean = EEGsum$meanVal, 
                                           sd = EEGsum$sdVal),
                            'poisson' = dpois(round(x), 
                                           lambda = EEGsum$meanVal), 
                            'logNorm' = dlnorm(x, 
                                               meanlog = EEGsum$logMean, 
                                               sdlog = EEGsum$logSD), 
                            'xVal' = x)


analyticDists <- analyticDists %>% pivot_longer(cols = names(analyticDists)[-4], 
                                               values_to = 'density', 
                                               names_to = 'dist')



df$dist = "empirical data"
df %>% ggplot(aes(x = pow_delta_O, color = dist)) + 
  geom_freqpoly(aes(y = ..density..), linewidth = 2, bins = 50) +
  geom_line(data = analyticDists, aes(x = xVal, y = density), linewidth = 2) + 
  brightCol +  myTheme + 
  ggtitle("EEG data") 



```

## make it into a function for ease of use

```{r functionitize}

plotAgainstDists <- function(df, varOfInterest, sumVals, myTitle, binCount = 50){

  x <- df[[varOfInterest]]
analyticDists <- data.frame(
                            'norm' = dnorm(x, 
                                           mean = sumVals$meanVal, 
                                           sd = sumVals$sdVal),
                            'poisson' = dpois(round(x), 
                                           lambda = sumVals$meanVal), 
                            'logNorm' = dlnorm(x + sumVals$offSet, 
                                               meanlog = sumVals$logMean, 
                                               sdlog = sumVals$logSD), 
                            'xVal' = x)


analyticDists <- analyticDists %>% pivot_longer(cols = names(analyticDists)[-4], 
                                               values_to = 'density', 
                                               names_to = 'dist')



df$dist = "empirical data"
df %>% ggplot(aes_string(x = varOfInterest, color = 'dist')) + 
  geom_freqpoly(aes(y = ..density..), linewidth = 2, bins = binCount) +
  geom_line(data = analyticDists, aes(x = xVal, y = density), linewidth = 2) + 
  brightCol +  myTheme + 
  ggtitle(myTitle) 

}


```

## my emails per day 

```{r email}

df = read.csv(paste0(gitHubPath, '\\data\\emailTimes.csv'))


df <- df %>% group_by(DayOfWeek, Month, DayOfMonth, Year) %>% 
  summarize(count = n(), .groups = 'drop')

# df = df %>% filter(DayOfWeek %in% c('Mon', 'Tue', 'Wed', 'Thu', 'Fri'))
# df = df %>% filter(DayOfWeek %in% c('Sat', 'Sun'))

df$dummy = as.factor(1) 

ggplot(df, aes(x = count, color = dummy, fill = dummy)) + 
  geom_histogram(breaks = seq(0, 51, 3)) + 
  myTheme + brightFil + brightCol + theme(legend.position = 'none')

```
## emails description

```{r descriptives for emails}

emailSum <- df %>% summarize(setName = 'emailsPerDay', 
                             meanVal = mean(count), 
                             sdVal = sd(count), 
                             logMean = mean(log(count)),
                             logSD = sd(log(count)), 
                             offSet = 0) 



```

## email frequency against analytic distributions 

```{r analytic comparison for email frequency}

plotAgainstDists(df, "count",  emailSum, 'email')


```

## sp 500 daily change 

```{r sp 500}

df = read.csv(paste0(gitHubPath, '\\data\\SP500HistoricalData.csv'))
names(df)[6] <- 'Change'

df$Change = unlist(lapply(df$Change, function(x) as.numeric(strsplit(x, '%')[[1]])))

df$dummy = as.factor(1) 

ggplot(df, aes(x = Change, color = dummy, fill = dummy)) +
  geom_histogram(breaks = seq(-2, 2, .1)) + 
  myTheme + brightFil + brightCol + theme(legend.position = 'none') + 
  xlab('Daily Change (%)') 

```

## sp 500 description

```{r descriptives for sp500}

sp500Sum <- df %>% summarize(setName = 'sp500', 
                               meanVal = mean(Change), 
                               sdVal = sd(Change), 
                               logMean = mean(log(Change-min(Change)+.01)),
                               logSD = sd(log(Change-min(Change)+.01)), 
                               offSet = -min(Change)+.01) 



```

## sp 500 changes against analytic distributions 

```{r analytic comparison for sp500}


plotAgainstDists(df, "Change", sp500Sum, 'sp500')


```

## human memory performance

```{r human memory}
df = read.csv(paste0(gitHubPath, 'data\\MemoryData.csv'))

df <- df %>% filter(TargALL_UVSD_d >=0)

#first plot the histogram of country total populations 
df$dummy = as.factor(1)
df %>% ggplot(aes(x = TargALL_UVSD_d, color = dummy, fill = dummy)) + 
  geom_histogram(bins = 10) + 
  myTheme + brightCol + brightFil + theme(legend.position = 'none')


```

## human memory description

```{r descriptives for human mem}

memSum <- df %>% summarize(setName = 'humMem', 
                               meanVal = mean(TargALL_UVSD_d), 
                               sdVal = sd(TargALL_UVSD_d),  
                               logMean = mean(log(TargALL_UVSD_d)),
                               logSD = sd(log(TargALL_UVSD_d)), 
                               offSet = 0) 



```

## human memory against analytic distributions 

```{r analytic comparison for human memory}

plotAgainstDists(df, "TargALL_UVSD_d",memSum, "mem data", 20)


```
