---
title: 'QQ plots'
Section: 'Determining Distribution'
Author: "Adam Dede"
Course: "Data to Decisions: Master Statistics with R and tidyverse"
Course_URL: "https://www.udemy.com/course/statsr_ad" 
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

gitHubPath <- file.path(path.expand("~"), "GitHub", 
                        "MasterStatsUsingR")


library(tidyverse) # data manipulation and visualization
source(file.path(gitHubPath, 'courseTheme.R'))
library(pracma)
# library(qqplotr)

```

## plotAgainstDists function from last video

```{r functionitize}
#you could experiment with putting this into another separate R script
#the same way that we did with the courseTheme! 


plotAgainstDists <- function(df, varOfInterest, sumVals, 
                             myTitle, binCount = 50){

  x <- df[[varOfInterest]]
  analyticDists <- data.frame(
                            'norm' = dnorm(x, 
                                           mean = sumVals$meanVal, 
                                           sd = sumVals$sdVal),
                            'poisson' = dpois(round(x), 
                                           lambda = sumVals$meanVal), 
                            'logNorm' = dlnorm(x + sumVals$offSet, 
                                           meanlog = sumVals$logMean, 
                                           sdlog = sumVals$logSD), 
                            'xVal' = x)


  analyticDists <- analyticDists %>% pivot_longer(
                        cols =  names(analyticDists)[-4], 
                        values_to = 'density', 
                        names_to = 'dist')



  df$dist = "empirical data"
  df %>% ggplot(aes_string(x = varOfInterest, color = 'dist')) + 
    geom_freqpoly(aes(y = after_stat(density)), 
                  linewidth = 2, bins = binCount) +
    geom_line(data = analyticDists, aes(x = xVal, y = density),
              linewidth = 2) + 
    brightCol +  myTheme + 
    ggtitle(myTitle) 

}


```

## human EEG data 

```{r EEG}
#load in the restingEEG dataset. 
#Key variable name: pow_delta_O
df = read.csv(file.path(gitHubPath, 'data',
                          'restingEEG.csv'))

df$dummy = as.factor(1) 
df <- df %>% filter(pow_delta_O < 250, pow_delta_O > 0)

ggplot(df, aes(x = pow_delta_O, color = dummy, fill = dummy)) + 
  geom_histogram() + 
  myTheme + brightFil + brightCol + theme(legend.position = 'none')

```

## human EEG description

```{r descriptives for EEG}

EEGsum <- df %>% summarize(setName = 'deltaPow', 
                               meanVal = mean(pow_delta_O),  
                               sdVal = sd(pow_delta_O), 
                               logMean = mean(log(pow_delta_O)),
                               logSD = sd(log(pow_delta_O)), 
                               offSet = 0) 



```

## human EEG QQ plots

```{r qq for EEG}
varOfInterest = "pow_delta_O"
sumVals = EEGsum
qVals = seq(.01, 1, .01)
plotAgainstDists(df, "pow_delta_O", 
                   EEGsum, 'EEG data')
  
  

qqVals = data.frame('empirical' = quantile(df[[varOfInterest]],
                                           qVals),
                    'norm'      = qnorm(qVals, 
                                    mean = sumVals$meanVal,
                                    sd = sumVals$sdVal), 
                    'logNorm'   = qlnorm(qVals, 
                                    meanlog = sumVals$logMean, 
                                    sdlog = sumVals$logSD), 
                    'poisson'   = qpois(qVals, 
                                    lambda = sumVals$meanVal), 
                    'unif'      = qunif(qVals, 
                                    min = min(df[[varOfInterest]]), 
                                    max = max(df[[varOfInterest]])),
                    'qVals'     = qVals)


ggplot(qqVals, aes(x = norm, y = empirical, color = qVals)) + 
  geom_point(size = 3) + 
  geom_abline(slope = 1, intercept = 0)

ggplot(qqVals, aes(x = logNorm, y = empirical, color = qVals)) + 
  geom_point(size = 3) + 
  geom_abline(slope = 1, intercept = 0)

ggplot(qqVals, aes(x = poisson, y = empirical, color = qVals)) + 
  geom_point(size = 3) + 
  geom_abline(slope = 1, intercept = 0)

ggplot(qqVals, aes(x = unif, y = empirical, color = qVals)) + 
  geom_point(size = 3) + 
  geom_abline(slope = 1, intercept = 0)

  



```

## QQ plot fuction

```{r qq function}
  
qqPlots <- function(df, varOfInterest, qVals, sumVals){
  qqVals = data.frame('empirical' = quantile(df[[varOfInterest]], 
                                    qVals),
                      'norm'      = qnorm(qVals, 
                                    mean = sumVals$meanVal, 
                                    sd = sumVals$sdVal), 
                      'logNorm'   = qlnorm(qVals, 
                                    meanlog = sumVals$logMean, 
                                    sdlog = sumVals$logSD), 
                      'poisson'   = qpois(qVals, 
                                    lambda = sumVals$meanVal), 
                      'unif'      = qunif(qVals, 
                                    min = min(df[[varOfInterest]]), 
                                    max = max(df[[varOfInterest]])),
                      'qVals'     = qVals)
    
  qqVals$logNorm <- qqVals$logNorm - sumVals$offSet
    
  out <- ggplot(qqVals, aes(x = norm, 
                            y = empirical, 
                            color = qVals)) + 
      geom_point(size = 3) + 
      geom_abline(slope = 1, intercept = 0)
    print(out)
    
  out <- ggplot(qqVals, aes(x = logNorm, 
                            y = empirical, 
                            color = qVals)) + 
      geom_point(size = 3) + 
      geom_abline(slope = 1, intercept = 0)
    print(out)
    
    out <- ggplot(qqVals, aes(x = poisson, 
                              y = empirical, 
                              color = qVals)) + 
      geom_point(size = 3) + 
      geom_abline(slope = 1, intercept = 0)
    print(out)
    
    out <- ggplot(qqVals, aes(x = unif, 
                              y = empirical, 
                              color = qVals)) + 
      geom_point(size = 3) + 
      geom_abline(slope = 1, intercept = 0)
    print(out)
  
}



```

## my emails per day 

```{r email}

df = read.csv(file.path(gitHubPath, 'data',
                      'emailTimes.csv'))


df <- df %>% group_by(DayOfWeek, Month, DayOfMonth, Year) %>%
  summarize(count = n(), .groups = 'drop')

df$dummy = as.factor(1) 

ggplot(df, aes(x = count, color = dummy, fill = dummy)) + 
  geom_histogram(breaks = seq(0, 51, 3)) + 
  myTheme + brightFil + brightCol + theme(legend.position = 'none')

```
## emails description

```{r descriptives for emails}

emailSum <- df %>% summarize(setName = 'emailsPerDay', 
                               meanVal = mean(count),
                               sdVal = sd(count), 
                               logMean = mean(log(count)),
                               logSD = sd(log(count)), 
                               offSet = 0) 



```

## email frequency against analytic distributions 

```{r analytic comparison for email frequency}

plotAgainstDists(df, "count", emailSum, 'email')
qqPlots(df, 'count', seq(.01, 1, .05), emailSum)

```

## sp 500 daily change 

```{r sp 500}

df = read.csv(file.path(gitHubPath, 'data', 
                        'SP500HistoricalData.csv'))
names(df)[6] <- 'Change'

df$Change = unlist(lapply(df$Change, function(x)
                  as.numeric(strsplit(x, '%')[[1]])))

df$dummy = as.factor(1) 

ggplot(df, aes(x = Change, color = dummy, fill = dummy)) +
  geom_histogram(breaks = seq(-2, 2, .1)) + 
  myTheme + brightFil + brightCol + 
  theme(legend.position = 'none') + 
  xlab('Daily Change (%)') 

```

## sp 500 description

```{r descriptives for sp500}

sp500Sum <- df %>% summarize(setName = 'sp500', 
                            meanVal = mean(Change), 
                            sdVal = sd(Change), 
        logMean = mean(log(Change-min(Change)+.01)),
        logSD = sd(log(Change-min(Change)+.01)), 
                               offSet = -min(Change)+.01) 



```

## sp 500 changes against analytic distributions 

```{r analytic comparison for sp500}


plotAgainstDists(df, "Change", 
                 sp500Sum, 'sp500')
qqPlots(df, 'Change', seq(.01, 1, .01), sp500Sum)

```

## human memory performance

```{r human memory}
df = read.csv(file.path(gitHubPath, 'data',
              'MemoryData.csv'))

df <- df %>% filter(TargALL_UVSD_d >=0)

#first plot the histogram of country total populations 
df$dummy = as.factor(1)
df %>% ggplot(aes(x = TargALL_UVSD_d, 
                  color = dummy, fill = dummy)) + 
  geom_histogram(bins = 10) + 
  myTheme + brightCol + brightFil + 
  theme(legend.position = 'none')


```

## human memory description

```{r descriptives for human mem}

memSum <- df %>% summarize(setName = 'humMem', 
                    meanVal = mean(TargALL_UVSD_d), 
                    sdVal = sd(TargALL_UVSD_d),  
                    logMean = mean(log(TargALL_UVSD_d)),
                    logSD = sd(log(TargALL_UVSD_d)), 
                    offSet = 0) 



```

## human memory against analytic distributions 

```{r analytic comparison for human memory}

plotAgainstDists(df, "TargALL_UVSD_d", 
                 memSum, "mem data", 20)
qqPlots(df, 'TargALL_UVSD_d', seq(.01, 1, .01), memSum)


```
