---
title: 'QQ plots'
author: "Adam Dede"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

gitHubPath = 'C:\\Users\\Adam\\Documents\\GitHub\\MasterStatsUsingR\\'

library(tidyverse) # data manipulation and visualization
source(paste0(gitHubPath, 'courseTheme.R'))
library(pracma)
# library(qqplotr)

```

## plotAgainstDists function from last video

```{r functionitize}
#you could experiment with putting this into another separate R script
#the same way that we did with the courseTheme! 
plotAgainstDists <- function(df, varOfInterest, empiricalBreaks, sumVals, title){

  L = length(empiricalBreaks)
  empiricalDensity = rep(0, L)
  
  for(ii in 1: L){
    empiricalDensity[ii] = sum(df[[varOfInterest]] < empiricalBreaks[ii])
  }
  empiricalDensity = (empiricalDensity[-1] - empiricalDensity[-L]) / 
                     (empiricalBreaks[-1] - empiricalBreaks[-L]) 
  empiricalBreaks = (empiricalBreaks[-1] + empiricalBreaks[-L]) / 2
  
  analyticDists <- data.frame('empiricalData' =empiricalDensity,
                              'norm' = dnorm(empiricalBreaks, 
                                             mean = sumVals$meanVal, 
                                             sd = sumVals$sdVal),
                              'poisson' = dpois(round(empiricalBreaks), 
                                             lambda = sumVals$meanVal), 
                              'logNorm' = dlnorm(empiricalBreaks+sumVals$offSet, 
                                                 meanlog = sumVals$logMean, 
                                                 sdlog = sumVals$logSD), 
                              'unif' = dunif(empiricalBreaks, 
                                             min = min(df[[varOfInterest]]), 
                                             max = max(df[[varOfInterest]])), 
                              'xVal' = empiricalBreaks)
  
  
  analyticDists <- analyticDists %>% pivot_longer(cols = names(analyticDists)[-6], 
                                                 values_to = 'density', 
                                                 names_to = 'dist')
  
  # analyticDists$xVal[analyticDists$dist == 'logNorm'] <- 
  #             analyticDists$xVal[analyticDists$dist == 'logNorm'] - sumVals$offSet
  
  analyticDists <- analyticDists %>% group_by(dist) %>%
                  mutate(density = density / sum(density))
  
  analyticDists %>% ggplot(aes(x = xVal, y = density, color = dist)) + 
    geom_line(linewidth = 2) + brightCol + myTheme + 
    ggtitle(title)

}



```

## human EEG data 

```{r EEG}
#load in the restingEEG dataset. There are a lot of variables in this dataset
#for this analysis let's just focus on the delta power in the occipital region 
#of the head. 
#Key variable name: pow_delta_O
df = read.csv(paste0(gitHubPath, '\\data\\restingEEG.csv'))

df$dummy = as.factor(1) 
df <- df %>% filter(pow_delta_O < 250, pow_delta_O > 0)

ggplot(df, aes(x = pow_delta_O, color = dummy, fill = dummy)) + 
  geom_histogram() + 
  myTheme + brightFil + brightCol + theme(legend.position = 'none')

```

## human EEG description

```{r descriptives for EEG}

EEGsum <- df %>% summarize(setName = 'deltaPow', 
                               meanVal = mean(pow_delta_O), 
                               medianVal = median(pow_delta_O), 
                               modeVal = Mode(round(pow_delta_O)), 
                               sdVal = sd(pow_delta_O), 
                               logMean = mean(log(pow_delta_O)),
                               logSD = sd(log(pow_delta_O)), 
                               offSet = 0, 
                               q25 = quantile(pow_delta_O, .25), 
                               q75 = quantile(pow_delta_O, .75), 
                               MADval = sum(abs(pow_delta_O - meanVal)) / 
                                          (n()-1)) 



```

## human EEG QQ plots

```{r qq for EEG}
  varOfInterest = "pow_delta_O"
  sumVals = EEGsum
  qVals = seq(.01, 1, .01)
  plotAgainstDists(df, "pow_delta_O", 
                   quantile(df$pow_delta_O, qVals), 
                   EEGsum, 'EEG data')
  
  

  qqVals = data.frame('empirical' = quantile(df[[varOfInterest]], qVals),
                      'norm'      = qnorm(qVals, mean = sumVals$meanVal, 
                                          sd = sumVals$sdVal), 
                      'logNorm'   = qlnorm(qVals, meanlog = sumVals$logMean, 
                                           sdlog = sumVals$logSD), 
                      'poisson'   = qpois(qVals, lambda = sumVals$meanVal), 
                      'unif'      = qunif(qVals, 
                                          min = min(df[[varOfInterest]]), 
                                          max = max(df[[varOfInterest]])),
                      'qVals'     = qVals)
  
  
  ggplot(qqVals, aes(x = norm, y = empirical, color = qVals)) + 
    geom_point(size = 3) + 
    geom_abline(slope = 1, intercept = 0)
  
  ggplot(qqVals, aes(x = logNorm, y = empirical, color = qVals)) + 
    geom_point(size = 3) + 
    geom_abline(slope = 1, intercept = 0)
  
  ggplot(qqVals, aes(x = poisson, y = empirical, color = qVals)) + 
    geom_point(size = 3) + 
    geom_abline(slope = 1, intercept = 0)
  
  ggplot(qqVals, aes(x = unif, y = empirical, color = qVals)) + 
    geom_point(size = 3) + 
    geom_abline(slope = 1, intercept = 0)
  
  



```

## QQ plot fuction

```{r qq function}
  
  qqPlots <- function(df, varOfInterest, qVals, sumVals){

    qqVals = data.frame('empirical' = quantile(df[[varOfInterest]], qVals),
                        'norm'      = qnorm(qVals, mean = sumVals$meanVal, 
                                            sd = sumVals$sdVal), 
                        'logNorm'   = qlnorm(qVals, meanlog = sumVals$logMean, 
                                             sdlog = sumVals$logSD), 
                        'poisson'   = qpois(qVals, lambda = sumVals$meanVal), 
                        'unif'      = qunif(qVals, 
                                            min = min(df[[varOfInterest]]), 
                                            max = max(df[[varOfInterest]])),
                        'qVals'     = qVals)
    
    qqVals$logNorm <- qqVals$logNorm - sumVals$offSet
    
    out <- ggplot(qqVals, aes(x = norm, y = empirical, color = qVals)) + 
      geom_point(size = 3) + 
      geom_abline(slope = 1, intercept = 0)
    print(out)
    
    out <- ggplot(qqVals, aes(x = logNorm, y = empirical, color = qVals)) + 
      geom_point(size = 3) + 
      geom_abline(slope = 1, intercept = 0)
    print(out)
    
    out <- ggplot(qqVals, aes(x = poisson, y = empirical, color = qVals)) + 
      geom_point(size = 3) + 
      geom_abline(slope = 1, intercept = 0)
    print(out)
    
    out <- ggplot(qqVals, aes(x = unif, y = empirical, color = qVals)) + 
      geom_point(size = 3) + 
      geom_abline(slope = 1, intercept = 0)
    print(out)
  
  }



```

## my emails per day 

```{r email}

df = read.csv(paste0(gitHubPath, '\\data\\emailTimes.csv'))


df <- df %>% group_by(DayOfWeek, Month, DayOfMonth, Year) %>% 
  summarize(count = n(), .groups = 'drop')

df$dummy = as.factor(1) 

ggplot(df, aes(x = count, color = dummy, fill = dummy)) + 
  geom_histogram(breaks = seq(0, 51, 3)) + 
  myTheme + brightFil + brightCol + theme(legend.position = 'none')

```
## emails description

```{r descriptives for emails}

emailSum <- df %>% summarize(setName = 'emailsPerDay', 
                               meanVal = mean(count), 
                               medianVal = median(count), 
                               modeVal = Mode(round(count)), 
                               sdVal = sd(count), 
                               logMean = mean(log(count)),
                               logSD = sd(log(count)), 
                               offSet = 0, 
                               q25 = quantile(count, .25), 
                               q75 = quantile(count, .75), 
                               MADval = sum(abs(count - meanVal)) / 
                                          (n()-1)) 



```

## email frequency against analytic distributions 

```{r analytic comparison for email frequency}

plotAgainstDists(df, "count", c(min(df$count):max(df$count)), emailSum, 'email')
qqPlots(df, 'count', seq(.01, 1, .05), emailSum)

```

## sp 500 daily change 

```{r sp 500}

df = read.csv(paste0(gitHubPath, '\\data\\SP500HistoricalData.csv'))
names(df)[6] <- 'Change'

df$Change = unlist(lapply(df$Change, function(x) as.numeric(strsplit(x, '%')[[1]])))

df$dummy = as.factor(1) 

ggplot(df, aes(x = Change, color = dummy, fill = dummy)) +
  geom_histogram(breaks = seq(-2, 2, .1)) + 
  myTheme + brightFil + brightCol + theme(legend.position = 'none') + 
  xlab('Daily Change (%)') 

```

## sp 500 description

```{r descriptives for sp500}

sp500Sum <- df %>% summarize(setName = 'sp500', 
                               meanVal = mean(Change), 
                               medianVal = median(Change), 
                               modeVal = Mode(round(Change)), 
                               sdVal = sd(Change), 
                               logMean = mean(log(Change-min(Change)+.001)),
                               logSD = sd(log(Change-min(Change)+.001)), 
                               offSet = -min(Change)+.001, 
                               q25 = quantile(Change, .25), 
                               q75 = quantile(Change, .75), 
                               MADval = sum(abs(Change - meanVal)) / 
                                          (n()-1)) 



```

## sp 500 changes against analytic distributions 

```{r analytic comparison for sp500}


plotAgainstDists(df, "Change", quantile(df$Change, seq(.01,1,.025)), 
                 sp500Sum, 'sp500')
qqPlots(df, 'Change', seq(.01, 1, .01), sp500Sum)

```

## human memory performance

```{r human memory}
df = read.csv(paste0(gitHubPath, 'data\\MemoryData.csv'))

df <- df %>% filter(TargALL_UVSD_d >=0)

#first plot the histogram of country total populations 
df$dummy = as.factor(1)
df %>% ggplot(aes(x = TargALL_UVSD_d, color = dummy, fill = dummy)) + 
  geom_histogram(bins = 10) + 
  myTheme + brightCol + brightFil + theme(legend.position = 'none')


```

## human memory description

```{r descriptives for human mem}

memSum <- df %>% summarize(setName = 'humMem', 
                               meanVal = mean(TargALL_UVSD_d), 
                               medianVal = median(TargALL_UVSD_d), 
                               modeVal = Mode(round(TargALL_UVSD_d)), 
                               sdVal = sd(TargALL_UVSD_d),  
                               logMean = mean(log(TargALL_UVSD_d)),
                               logSD = sd(log(TargALL_UVSD_d)), 
                               offSet = 0, 
                               q25 = quantile(TargALL_UVSD_d, .25), 
                               q75 = quantile(TargALL_UVSD_d, .75), 
                               MADval = sum(abs(TargALL_UVSD_d - meanVal)) / 
                                          (n()-1)) 



```

## human memory against analytic distributions 

```{r analytic comparison for human memory}

plotAgainstDists(df, "TargALL_UVSD_d", 
                 quantile(df$TargALL_UVSD_d, seq(.01,1,.025)), 
                 memSum, "mem data")
qqPlots(df, 'TargALL_UVSD_d', seq(.01, 1, .01), memSum)


```
