---
Title: 'Combining Data from multiple Sources'
Section: 'Tidyverse'
Author: "Adam Dede"
Course: "Data to Decisions: Master Statistics with R and tidyverse"
Course_URL: "..." 
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---

#### Examining the relationship between waste management and GDP

Data can be found here: 
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-05-21


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
myGitHubPath = 'C:\\Users\\Adam\\Documents\\GitHub\\MasterStatsUsingR\\'
# Load packages
library(tidyverse) # data manipulation and visualization
library(janitor) #helps to rename titles
library(tidytuesdayR) #package for loading tidy tuesday datasets 
```

## Load Data in three CSVs and use janitor to make names R appropriate

```{r load data}

# Load the data 
#coastal vs mismanaged waste
cw <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/coastal-population-vs-mismanaged-plastic.csv")

#mismanaged waste vs gdp
mg <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/per-capita-mismanaged-plastic-waste-vs-gdp-per-capita.csv")

#total waste vs. gdp
wg <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/per-capita-plastic-waste-vs-gdp-per-capita.csv")

# inspect the data 
head(cw)
head(mg)
head(wg)
names(cw)
cw <- janitor::clean_names(cw)
names(cw)
names(mg)
mg <- janitor::clean_names(mg)
names(mg)
names(wg)
wg <- janitor::clean_names(wg)
names(wg)



```

## Hard code easier to deal with names for all files

```{r hardcode rename}

names(cw)
colnames(cw) <- c('country', 'countryCode', 'year', 'mismanagedTotal',
                  'coastalPop', 'totalPop')
names(cw)
names(mg)
colnames(mg) <- c('country', 'countryCode', 'year', 'mismanagedPerCap',
                  'GDPperCap', 'totalPop')
names(mg)
names(wg)
colnames(wg) <- c('country', 'countryCode', 'year', 'totWastePerCap',
                  'gdpPerCap2', 'totalPop')
names(wg)


```

## IS the year 2010 going to be a good year to analyze for all files? 

```{r examine valid data}

# summarize how many good observations there are for each variable of interest

#key variable in coast_vs_waste is the mismanaged plastic
dataCount_cw <- cw %>%
  select(year, mismanagedTotal) %>%
  group_by(year) %>%
  summarize(naCount = sum(is.na(mismanagedTotal)),
            n = n()) %>% 
  mutate(goodVals = n - naCount)

#key variable in mismanaged_vs_gdp is per capita mismanaged plastic
dataCount_mg <- mg %>%
  select(year, mismanagedPerCap) %>%
  group_by(year) %>%
  summarize(naCount = sum(is.na(mismanagedPerCap)),
    n = n()) %>% 
  mutate(goodVals = n - naCount)

#key variable in waste_vs_gdp is total per capital plastic waste
dataCount_wg <- wg %>%
  select(year, totWastePerCap) %>%
  group_by(year) %>%
  summarize(naCount = sum(is.na(totWastePerCap)), 
    n = n()) %>% 
  mutate(goodVals = n - naCount)


dataCount_cw %>% ggplot(aes(x = year-.5, y = goodVals)) +
  geom_point(color = 'green') + 
  geom_point(data = dataCount_mg, aes(x = year, y = goodVals), color = 'red')+
  geom_point(data = dataCount_wg, aes(x = year+.5, y = goodVals), 
             color = 'black') +
  xlim(c(2000, 2015))


```

## Filtering down to only the year 2010 and getting rid of NA values 

```{r get rid of NA and select year}
#down select to remove missing and check that we only have one year
cw = cw %>% filter(!is.na(mismanagedTotal))
mg <- mg %>% filter(!is.na(mismanagedPerCap))
wg <- wg %>% filter(!is.na(totWastePerCap))

#check that there's only one year for each country in each file: 
countryCounts = cw %>%
  group_by(country) %>%
  summarize(n = n())

max(countryCounts$n)

countryCounts = mg %>%
  group_by(country) %>%
  summarize(n = n())

max(countryCounts$n)

countryCounts = wg %>%
  group_by(country) %>%
  summarize(n = n())

max(countryCounts$n)

```

## Merge the three data frames together into one large dataframe!

```{r merge}
names(cw)
names(mg)
df <- merge(cw, mg, by = "countryCode", suffixes = c("_df1", "_df2"))
names(df)
df <- df %>%
  select(-ends_with("_df2"))
names(df)
new_column_names <- gsub("_df1$", "", names(df))
names(df) <- new_column_names
names(df)

df <- merge(df, wg, by = "countryCode", suffixes = c("_df1", "_df2"))
df <- df %>%
  select(-ends_with("_df2"))
new_column_names <- gsub("_df1$", "", names(df))
names(df) <- new_column_names
```

## Remove incomplete rows! 

```{r remove incomplete}

df <- df %>% drop_na()

```

## save our cleaned data frame for later use! 
```{r save the clean .csv}
write.csv(df, paste0(myGitHubPath, 'data\\plasticWaste.csv'))

```
# Data Visualisation: Histograms 

Create histograms to display the distributions of all variables in the data

```{r histograms}

df %>% ggplot(aes(x = mismanagedTotal))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = coastalPop))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = totalPop))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = mismanagedPerCap))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = gdpPerCap))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = totWastePerCap))+
  geom_histogram() +
theme_classic()


```
