---
title: 'Central Tendency'
Section: 'Simulating Data from Distributions'
Author: "Adam Dede"
Course: "Data to Decisions: Master Statistics with R and tidyverse"
Course_URL: "https://www.udemy.com/course/statsr_ad" 
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)


gitHubPath <- file.path(path.expand("~"), "GitHub", 
                        "MasterStatsUsingR")


library(tidyverse) # data manipulation and visualization
source(file.path(gitHubPath, 'courseTheme.R'))
library(pracma) #practical numerical math functions 
#(needed for Mode)

```

## simulate data from all four distributions we know

```{r simulate data}

sampSize = 10000

df = data.frame('norm' = rnorm(sampSize, 3, 1),
                'lnorm' = rlnorm(sampSize,log(3), 1), 
                'unif' = runif(sampSize,-2, 8),
                'pois' = rpois(sampSize, lambda = 3))

df <- df %>% pivot_longer(cols = c(names(df)), 
                   values_to = 'observed', 
                   names_to = 'dist')


```
## let's work with just the normal distribution data for now, start with plotting
```{r plot norm dist}

p <- df %>% filter(dist == 'norm') %>%
  ggplot(aes(x = observed, group = dist, color = dist)) + 
  geom_freqpoly(breaks = seq(-1, 7, .1), linewidth = 1) + 
  myTheme + brightCol 

print(p) 
```
## calculate the central tendancy measures for the normal distribution

``` {r norm analysis}

normDat = df$observed[df$dist == 'norm'] 

#what's the mean value? 
# mean = sum(x) / length(x) 
sum(normDat) / length(normDat) -> meanVal
  
#what's the median value? 
#value in the middle: 
normDat = sort(normDat)
if(length(normDat) %% 2 ==0) { #even length! 
  midVal = length(normDat)/2
  (normDat[midVal] + normDat[midVal+1]) / 2 -> medVal
} else { #odd length!
  midVal = length(normDat)/2  + .5 
  normDat[midVal] -> medVal
}
  
#what's the modal value? 
#this line is a bit of a hack to get a central mode in continuous data: 
round(normDat, 1) -> normDat  

uniSet = unique(normDat)
#are there even any repeated values? 
if(length(uniSet) == length(normDat)){ #every value unique! 
  sort(normDat)[1] -> modVal
} else { #there's going to be a proper mode! 
  lengths = rep(0, length(uniSet))
  
  for(jj in 1:length(uniSet)){
    lengths[jj] = sum(normDat==uniSet[jj])
  }
  
  uniSet[which.max(lengths)] -> modVal
}

# Add vertical lines for central tendency measures
p + geom_vline(xintercept = meanVal, linetype = 'solid', size = 3, color = 'red') + 
    geom_vline(xintercept = modVal, linetype = 'dashed', size = 3, color = 'red') + 
    geom_vline(xintercept = medVal, linetype = 'dotted', size = 3, color = 'red')


```

## calculate the central tendancy measures for the different distributions
``` {r central tendency measures}

#how many distributions do we have data for? 
dists = unique(df$dist)
L = length(dists)

#create a data.frame to put all the central tendancy values into: 
group_averages = data.frame('dist' = dists, 
                            'meanVal' = rep(NA,L), 
                            'modeVal' = rep(NA,L), 
                            'medianVal' = rep(NA,L))

#Challenge: can you write a for loop to fill out the group_averages data frame? 

for(ii in 1:L){
  curDat = df$observed[df$dist == dists[ii]]
  
  #what's the mean value? 
  # mean = sum(x) / length(x) 
  group_averages$meanVal[ii] = sum(curDat) / length(curDat)
  
  #what's the median value? 
  #value in the middle: 
  curDat = sort(curDat)
  if(length(curDat) %% 2 ==0) { #even length! 
    midVal = length(curDat)/2
    group_averages$medianVal[ii] = (curDat[midVal] + curDat[midVal+1]) / 2
  } else { #odd length!
    midVal = length(curDat)/2  + .5 
    group_averages$medianVal[ii] = curDat[midVal]
  }
  
  #what's the modal value? 
  round(curDat, 1) -> curDat
  uniSet = unique(curDat)
  #are there even any repeated values? 
  if(length(uniSet) == length(curDat)){ 
    #every value unique! 
    group_averages$modeVal[ii] = curDat[1]
  } else { #there's going to be a proper mode! 
    lengths = rep(0, length(uniSet))
    
    for(jj in 1:length(uniSet)){
      lengths[jj] = sum(curDat==uniSet[jj])
    }
    
    group_averages$modeVal[ii] = uniSet[which.max(lengths)]
    }
  
}

group_averages

group_averages <- group_averages %>% pivot_longer(cols = 
                                c('meanVal', 'modeVal', 'medianVal'), 
                                names_to = 'measure', 
                                values_to = 'centralTend') 


```

## calculating the central tendency using builtin functions!

```{r builtin demo}
# Calculate group means
df  %>%
  group_by(dist) %>%
  summarise(mean_val = mean(observed),
            mode_val = Mode(round(observed,1)),
            median_val = median(observed)) -> builtinAvg 

```
## plot for the poisson distribution data with central tendency values

```{r pois plot}

p <- df %>% filter(dist == 'pois') %>%
  ggplot(aes(x = observed, group = dist, color = dist, fill = dist)) + 
  geom_freqpoly(breaks = seq(-1.5, 9.5, 1), linewidth = 3) + 
  myTheme + brightCol + brightFil 

print(p)

#for now, just hard code in the central tendency values 
p + geom_vline(data = filter(group_averages, dist == 'pois'),
               aes(xintercept = centralTend, linetype = measure), 
               size = 3, color = 'red') 


```
## can you make plots for the uniform and log normal distributions? 

```{r plot the other distributions}

#lnorm: 
p <- df %>% filter(dist == 'lnorm') %>%
  ggplot(aes(x = observed, group = dist, color = dist, fill = dist)) + 
  geom_freqpoly(breaks = seq(-1.5, 9.5, 1), linewidth = 3) + 
  myTheme + brightCol + brightFil 

print(p)

#for now, just hard code in the central tendency values 
p + geom_vline(data = filter(group_averages, dist == 'lnorm'),
               aes(xintercept = centralTend, linetype = measure), 
               size = 3, color = 'red') 
  
#unif: 
p <- df %>% filter(dist == 'unif') %>%
  ggplot(aes(x = observed, group = dist, color = dist, fill = dist)) + 
  geom_freqpoly(breaks = seq(-1.5, 9.5, 1), linewidth = 3) + 
  myTheme + brightCol + brightFil 

print(p)

#for now, just hard code in the central tendency values 
p + geom_vline(data = filter(group_averages, dist == 'unif'),
               aes(xintercept = centralTend, linetype = measure), 
               size = 3, color = 'red') 


```
## make all plots in one go? merge the central tendencies with the data to make it happen! 

```{r make all plots}
## challenge: 
## 1) Calculate central tendency with builtin functions
## 2) generate all the freq_poly plots for all distributions at once! 
## 3) put all the central tendency measures on all of them at once! 
## hint: use faceting to make the plots easy to read

#calcualte central tendency: 
df  %>%
  group_by(dist) %>%
  summarise(mean_val = mean(observed),
            mode_val = Mode(round(observed,1)),
            median_val = median(observed)) -> builtinAvg 

#pivot longer to facilitate plotting
builtinAvg <- builtinAvg %>% 
              pivot_longer(cols = -c('dist'), 
                           names_to = 'type', 
                           values_to = 'centralTend')

df %>% ggplot(aes(x = observed, color = dist)) + 
  geom_freqpoly(breaks = seq(-1.5, 8, 1), linewidth = 2) + 
  facet_grid(dist~.) + myTheme + brightCol -> p

p + geom_vline(data = builtinAvg,
               aes(xintercept = centralTend, linetype = type), 
               size = 3, color = 'red') +
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank())

#challenge question: If you regenerate the simulated data, 
#                    what will happen to the mode of the 
#                    uniform data? 


```

