---
Title: 'Probability of a Sample'
Section: 'Z-test'
Author: "Adam Dede"
Course: "Data to Decisions: Master Statistics with R and tidyverse"
Course_URL: "..." 
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

gitHubPath = 'C:\\Users\\Adam\\Documents\\GitHub\\MasterStatsUsingR\\'

library(tidyverse) # data manipulation and visualization
source(paste0(gitHubPath, 'courseTheme.R'))
library(pracma)

```

## heights of basketball players 

```{r data}
#basketball player heights
df = data.frame('heights' = c(72,74,76,80,78,73,73,71,78,81,75,75,79,77,79))

#overall distribution of heights statistics: 
M = 70
S = 4

```

## Visualize players against the overall population 

```{r visualization}


dfOverall = data.frame('norm' = dnorm(seq(50,90,1), M, S))
dfOverall$dist = 'overallPop'

df$dist = 'NBA'

df %>% ggplot(aes(x = heights, color = dist)) + 
  geom_freqpoly(aes(y = ..density..), linewidth = 2, bins = 30) +
  geom_line(data = dfOverall, aes(x = seq(50,90,1), y = norm), linewidth = 2) + 
  brightCol +  myTheme 


```

## what is the probability of each basketball player height?  

```{r individual probs}

#z-score players' heights
#BUT z-score them relative to the overall distribution! 
df$zscore = (df$heights - M) / S

#What is the probability of each player's height? 
df$prob = 1-pnorm(df$zscore)

#it looks like they're all rare, but they're not all THAT rare
df %>% ggplot(aes(x = prob, color = dist, fill = dist)) + 
  geom_histogram() + myTheme + brightCol + brightFil



```

## What is the probability of this whole sample though   

```{r sample probability }

#let's use a little m to denote the sample mean as opposed to the population M
m = mean(df$heights)

#What is the probability of the mean height of our sample? 
meanProb = 1-pnorm(m, M, S)

#what if we use the multiplication rule? 
multiProb = prod(df$prob)



```

## simulate drawing height samples   

```{r simulate samples}
samples = 1000
sampSizes = c(1,2,3,4,5,10,15)
simResults = data.frame(meanVal = rep(0,samples*length(sampSizes)),
                        sampSize = rep(0,samples*length(sampSizes)))
ii = 1
for(sampSize in sampSizes){
  print(sampSize)
  for(samp in 1:samples){
    simResults$meanVal[ii] = mean(rnorm(sampSize, M,S))
    simResults$sampSize[ii] = sampSize
    ii = ii+1
  }
}
#what proportion of sample means from the overall distribution are greater
#than the sample mean from our NBA player sample? 
outSum = simResults %>% group_by(sampSize) %>%
  summarize(prop = sum(meanVal>m)/samples)

outSum$dummy = as.factor(1)
outSum %>% ggplot(aes(x = sampSize, y = prop, color = dummy)) + 
  geom_line(linewidth = 2) + myTheme + brightCol + 
  theme(legend.position = 'none') + 
  geom_hline(yintercept = meanProb, color = 'red') + 
  geom_hline(yintercept = multiProb, color = 'green')


```

## looking at the simulation from a different angle: what is the distribution of n=15 sample means?    

```{r simulation 2}
simResults$dummy = as.factor(1)


#when the sample size is 15, we never ever observe a sample MEAN as extreme
#as our NBA sample
simResults %>% filter(sampSize == 15) %>% 
  ggplot(aes(x=meanVal, color = dummy, fill = dummy)) + 
  geom_histogram() + myTheme + brightCol + brightFil + 
  geom_vline(xintercept = m, color = 'red', linewidth = 2)
  
#if the sample size is only 1 person, then our NBA sample mean is not so 
#extreme
simResults %>% filter(sampSize == 1) %>% 
  ggplot(aes(x=meanVal, color = dummy, fill = dummy)) + 
  geom_histogram() + myTheme + brightCol + brightFil + 
  geom_vline(xintercept = m, color = 'red', linewidth = 2)



```

## What's happening is that the variability of sample means reduces with sample size!  

```{r simulation 3}

outSum = simResults %>% group_by(sampSize) %>%
  summarize(prop = sum(meanVal>m)/samples, 
            sdVal = sd(meanVal), 
            meanSampleMean = mean(meanVal))


outSum$dummy = as.factor(1)
outSum %>% ggplot(aes(x = sampSize, y = sdVal, color = dummy)) + 
  geom_line(linewidth = 2) + myTheme + brightCol + ylim(c(0,4))

 
outSum %>% ggplot(aes(x = sampSize, y = meanSampleMean, color = dummy)) + 
  geom_line(linewidth = 2) + myTheme + brightCol + ylim(c(50,80))



```

## So, what is the probability of our sample of NBA players? 

```{r final prob}

#let's use the sd from our n=15 simulation 
1 - pnorm(m, M, outSum$sdVal[outSum$sampSize==15])

multiProb
meanProb


```

## scratch 

```{r all samplingDists}

simResults %>% ggplot(aes(x = meanVal, 
                          group = as.factor(sampSize), 
                          color = sampSize)) + 
  geom_freqpoly(aes(y = ..density..), linewidth = 2) + myTheme

```