---
title: 'Plastic Waste Project'
author: "Adam Dede"
date: "2024-01-02"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---

#### Examining the relationship between waste management and GDP

Original data can be found here: 
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-05-21

more information about the data can be found here: 
https://ourworldindata.org/plastic-pollution
https://www.science.org/doi/10.1126/sciadv.aaz5803#sec-4


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# Load packages
courseThemePath = 
  "G:\\My Drive\\Teaching\\stats\\RCourse\\t_test_vs_regression_analysis\\courseTheme.R"

library(tidyverse) # data manipulation and visualization
source(courseThemePath)
```

# load data that we previously cleaned and organized
```{r load plastic waste data}

df = read.csv('G:\\My Drive\\Teaching\\stats\\RCourse\\cleanData\\plasticWaste.csv')


```

# visualize where the median income is

calculate the median and plot it on a histogram of income

```{r rich/poor median}



```

# split the countries at the median income and do the t-test comparing these groups

```{r t-test on rich/poor}

```

# visualization using box plot

```{r boxplot}



```

# visualization using scatter plot

```{r scatter}



```

## do a linear regression to predict mismanagedPerCap as a function of gdpPerCap

```{r linear regression}


```

## add a trend line to the scatter plot
```{r add scatter plot rend}

 



```

## examine the results of both analyses at once

```{r adding the mean values to the plot}



```

## residuals plot

```{r residuals plotting}



```

## calcualte R squared

```{r R squared calculations}




```

