---
title: 'Plastic Waste Project'
author: "Adam Dede"
date: "2024-01-02"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---

#### Examining the relationship between waste management and GDP

Original data can be found here: 
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-05-21

more information about the data can be found here: 
https://ourworldindata.org/plastic-pollution
https://www.science.org/doi/10.1126/sciadv.aaz5803#sec-4


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

gitHubPath = 'C:\\Users\\Adam\\Documents\\GitHub\\MasterStatsUsingR\\'

library(tidyverse) # data manipulation and visualization
source(paste0(gitHubPath, 'courseTheme.R'))
```

# load data that we previously cleaned and organized
```{r load plastic waste data}

df = read.csv('G:\\My Drive\\Teaching\\stats\\RCourse\\cleanData\\plasticWaste.csv')


```

# visualize where the median income is

calculate the median and plot it on a histogram of income

```{r rich/poor median}
medianGDP = median(df$gdpPerCap)

df %>% ggplot(aes(x = gdpPerCap, color = as.factor(year), 
                  fill = as.factor(year))) + 
  geom_histogram() + myTheme + brightCol + brightFil + 
  geom_vline(xintercept = medianGDP, linewidth = 3, color = 'green') + 
  theme(legend.position = 'none')



```

# split the countries at the median income and do the t-test comparing these groups

```{r t-test on rich/poor}
richDF <- df %>% filter(gdpPerCap >= medianGDP)
poorDF <- df %>% filter(gdpPerCap < medianGDP)
t.test(richDF$mismanagedPerCap, poorDF$mismanagedPerCap) 

```

# visualization using box plot

```{r boxplot}
df <- df %>% mutate(richPoor = gdpPerCap > medianGDP)
df$richPoor[df$richPoor] = 'rich'
df$richPoor[df$richPoor != 'rich'] = 'poor' 

df %>% ggplot(aes(group = richPoor, fill = richPoor, y = mismanagedPerCap)) + 
  geom_boxplot(color = 'white') +
  myTheme + brightCol + brightFil




```

# visualization using scatter plot

```{r scatter}
df %>% ggplot(aes(group = richPoor, color = richPoor, x = gdpPerCap, 
                  y = mismanagedPerCap)) + 
  geom_point(size = 5) +
  myTheme + brightCol + brightFil


```

## do a linear regression to predict mismanagedPerCap as a function of gdpPerCap

```{r linear regression}
richPoor.lm = lm(mismanagedPerCap ~ gdpPerCap, data = df)

summary(richPoor.lm)

```

## add a trend line to the scatter plot
```{r add scatter plot rend}
df %>% ggplot(aes(group = richPoor, color = richPoor, x = gdpPerCap, 
                  y = mismanagedPerCap)) + 
  geom_point(size = 5) +
  myTheme + brightCol + brightFil + 
  geom_smooth(method = "lm", color = 'red', se = FALSE) 

df %>% ggplot(aes(color = richPoor, x = gdpPerCap, 
                  y = mismanagedPerCap)) + 
  geom_point(size = 5) +
  myTheme + brightCol + brightFil + 
  geom_smooth(method = "lm", color = 'red', se = FALSE) 

```

## examine the results of both analyses at once

```{r adding the mean values to the plot}
temp <- df %>% 
  group_by(richPoor) %>%
  summarize(meanVal = mean(mismanagedPerCap))

df$groupMean = temp$meanVal[match(df$richPoor, temp$richPoor)]

df$linReg_pred = richPoor.lm$fitted.values

df %>% ggplot(aes(color = richPoor, x = gdpPerCap, 
                  y = mismanagedPerCap)) + 
  geom_point(size = 5) +
  myTheme + brightCol + brightFil + 
  geom_line(aes(x = gdpPerCap, y = groupMean), color = 'red') + 
  geom_line(aes(x = gdpPerCap, y = linReg_pred), color = 'magenta')


```

## residuals plot

```{r residuals plotting}
df$ttest_res = df$mismanagedPerCap - df$groupMean
df$linReg_res = df$mismanagedPerCap - df$linReg_pred

df %>% ggplot(aes(color = richPoor, x = gdpPerCap, 
                  y = ttest_res)) + 
  geom_point(size = 5) +
  myTheme + brightCol + brightFil

df %>% ggplot(aes(color = richPoor, x = gdpPerCap, 
                  y = linReg_res)) + 
  geom_point(size = 5) +
  myTheme + brightCol + brightFil


```

## calcualte R squared

```{r R squared calculations}
summary(richPoor.lm)
SStot = sum((df$mismanagedPerCap - mean(df$mismanagedPerCap))^2)
SSres = sum(df$linReg_res^2)
Rsq_linReg = 1 - SSres / SStot

SSres = sum(df$ttest_res^2)
Rsq_ttest = 1 - SSres / SStot

print(paste('Rsq lin reg: ', round(Rsq_linReg, 2), 
            ' Rsq ttest: ', round(Rsq_ttest, 2)))

richPoorttest.lm = lm(mismanagedPerCap ~ richPoor, data = df)
summary(richPoorttest.lm)



```

