---
title: 'Plastic Waste Project'
author: "Adam Dede"
date: "2024-01-02"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---

#### Examining the relationship between waste management and GDP

Original data can be found here: 
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-05-21

more information about the data can be found here: 
https://ourworldindata.org/plastic-pollution
https://www.science.org/doi/10.1126/sciadv.aaz5803#sec-4


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# Load packages
courseThemePath = 
  "G:\\My Drive\\Teaching\\stats\\RCourse\\t_test_vs_regression_analysis\\courseTheme.R"

library(tidyverse) # data manipulation and visualization
source(courseThemePath)
```

# load data that we previously cleaned and organized
```{r load plastic waste data}

df = read.csv('G:\\My Drive\\Teaching\\stats\\RCourse\\cleanData\\plasticWaste.csv')


```

# visualize where the median income is

calculate the median and plot it on a histogram of income

```{r rich/poor median}
medianGDP = median(df$gdpPerCap)

df %>% ggplot(aes(x =         ) + 
  geom_histogram() 


```

# split the countries at the median income and do the t-test comparing these groups

```{r t-test on rich/poor}
richDF <- df %>% 
poorDF <- df %>% 
t.test(richDF$mismanagedPerCap, poorDF$mismanagedPerCap) 

```

# visualization using box plot

```{r boxplot}
df <- df %>% mutate(richPoor =  )


df %>% ggplot(aes(group =     , fill =    , y =      )) + 
  geom_boxplot() 




```

# visualization using scatter plot

```{r scatter}
df %>% ggplot(aes(group =   , color =      , x =      , 
                  y =      )) + 
  geom_point( )  


```

## do a linear regression to predict mismanagedPerCap as a function of gdpPerCap

```{r linear regression}
richPoor.lm = lm(mismanagedPerCap ~   , data =   )



```

## add a trend line to the scatter plot
```{r add scatter plot rend}

  geom_smooth(method = "lm") 



```

## examine the results of both analyses at once

```{r adding the mean values to the plot}
temp <- df %>% 
  group_by(  ) %>%
  summarize(meanVal =  )

df$groupMean = temp$meanVal[match(  )]

df$linReg_pred =  

df %>% ggplot(aes(color =   , x =   , 
                  y =   )) + 
  geom_point( )  +
  geom_line(aes(x = gdpPerCap, y =  ), color = 'red') + 
  geom_line(aes(x = gdpPerCap, y =  ), color = 'magenta')


```

## residuals plot

```{r residuals plotting}
df$ttest_res = df$mismanagedPerCap - 
df$linReg_res = df$mismanagedPerCap - 

df %>% ggplot(aes(color = richPoor, x = , 
                  y = ttest_res)) + 
  geom_point( )  

df %>% ggplot(aes(color = richPoor, x = , 
                  y = linReg_res)) + 
  geom_point( )  


```

## calcualte R squared

```{r R squared calculations}
summary(richPoor.lm)
SStot =  
SSres =  
Rsq_linReg = 1 -  

SSres =  
Rsq_ttest = 1 -  

print(paste('Rsq lin reg: ', round(Rsq_linReg, 2), 
            ' Rsq ttest: ', round(Rsq_ttest, 2)))

richPoorttest.lm = lm(mismanagedPerCap ~ richPoor, data = df)
summary(richPoorttest.lm)



```

