---
title: 'Arrays and Matrices'
author: "Adam Dede"
date: "2024-03-19"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)



```

## creating and indexing in matrices works like vectors but with two dimensions

```{r make a matrix}


```
## switching to row major orientation
```{r populating matrices along rows}




```

## you can do math operations on whole matrices like on vectors with some extras
```{r matrix math}




```

## more indexing tricks 
```{r more matrix indexing tricks}




```
## using the dim function on matrices
```{r dim}




```
## matrices are vectors with row and column attributes
```{r matrix vector relationship}


```
## matrix indexing understanding check
```{r index challenge}



```
## matrix indexing understanding check
```{r index challenge}



```
## adding rows and columns to matrices
```{r rows and columns}


```
## recycling in matrices
```{r recycling}

```
##dimensions can get dropped and break your code
```{r dimension dropping}

```

## arrays (matrices with more than 2 dimensions)
```{r arrays}

```
