---
title: 'Arrays and Matrices'
author: "Adam Dede"
date: "2024-03-19"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)



```

## lists let you combine different data modes

```{r make a list}

#often, we want to combine data modes
x <- list(name='adam', age = 33, colGrad = T)
mode(x)
attributes(x)
str(x)
names(x)

```
## lists don't have to have names
```{r nameless list}

xNameless <- list('adam', 33, T)

#as with vectors, we can guess how to index
#by looking at a variable carefully
xNameless
xNameless[[1]]
x
x$name

#named and unnamed list style indexing combined:
x[['name']]


```

## we can also add names to an unnamed list
```{r adding names}

names(x)

xNameless
names(xNameless) <- c("name", "age", "colGrad")
xNameless

unname(xNameless)

```

## what's with the double square brackets? 
```{r double versus single brackets}

#a fourth and fifth way to index into lists: 
x['name']
x[1]

#what's the mode of double versus single bracketed indexing? 
mode(x[['name']])
mode(x['name'])

#Why do you think this doesn't work: 
x[[1:2]]
#but this does: 
x[1:2]

#lists are really just vectors that allow 
#anything to be at each index in mode 'list'

#double bracketing converts a single index
#back to its atomic mode and returns it.
#Single bracketing returns a new vector 
#that's still in mode 'list'


```
## adding to an already created list
```{r dim}

#adding to already created lists
x
x$height <- 175
x
x[5:7] <- c(T,F,T)
x
#deleting from a list
x[4] <- NULL
x
#how did the indices of items after 4 change?

#don't confuse NULL with NA or NAN
x[5] <- NA
x

#check the length of a list
length(x)


```
## matrices are vectors with row and column attributes
```{r matrix vector relationship}
#R thinks of matrices as vectors
#that just have row and column attributes
attributes(z)
length(z)
z[4]
#you can also flatten a matrix into a vector
z = c(z)
z[4]

```
## converting to a vector, but watch out for coercion! 
```{r unlist}

unlist(x) -> y
y

```
## lists of lists are kind of like what matrices are to vectors
```{r lists of lists}
x <- list(name='adam', age = 33, colGrad = T)
y <- list(name='sara', age = 35, colGrad = T)
z <- list(name='bob' , age = 21, colGrad = F)

allDat <- list(x,y,z)
#use names to set the names of items in top list
names(allDat) <- c('a', 'b', 'c')

```
## indexing challenge! 
```{r list of lists indexing}

#can you write as many ways as possible
#to find bob's age? Hint: find at least 8

allDat[[3]]$age


allDat[[3]][[2]]
allDat[[3]]$age
allDat[3]$age
allDat[3][[1]]$age
allDat[[3]]['age']
allDat[[3]][2]
allDat[['c']][[2]]
allDat[['c']]$age
allDat['c']$age
allDat['c'][[1]]$age
allDat[[3]]['age']
allDat[[3]][2]
allDat['c']$c$age
allDat$c$age
```
