---
title: 'Adding in Interactions'
Section: 'Multiple Linear Regression'
Author: "Adam Dede"
Course: "Data to Decisions: Master Statistics with R and tidyverse"
Course_URL: "..." 
output: 
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## Set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

gitHubPath = 'C:\\Users\\Adam\\Documents\\GitHub\\MasterStatsUsingR\\'

library(tidyverse) # data manipulation and visualization
source(paste0(gitHubPath, 'courseTheme.R'))
library(car)
# Load package
library(corrplot)

```

## Import tidy data on cardiac risk factors across the US and look at full model

```{r import}

df = read.csv(paste0(gitHubPath, '\\data\\cardioData_sum.csv'))

 

dfCut = df %>% select(-YearStart, -X, -LocationDesc)

predNames <- names(dfCut) %>%
  setdiff(c("LocationAbbr", "cardioDeath"))
# from last time, lostTeeth, noVeg, and smokeBan were our three best predictors
outputVars = c('lostTeeth', 'noVeg', 'smokeBan')
# remove these from the list of predictors 
predNames = predNames[!predNames %in% outputVars]
formula <- reformulate(outputVars, response = "cardioDeath")
mainOnlyMod = lm(formula, dfCut)
#remind ourselves of our model
summary(mainOnlyMod)


```

## Searching for a fourth variable found nothing, but are any marginal? 

```{r fourth}
check = T
possibleVars = c()
while(check){
  #repeat to add to model: 
  etaVals = rep(0, length(predNames))
  names(etaVals) = predNames
  #loop over predictors and pull Rsq values 
  for(varName in predNames){
    
    formula <- reformulate(c(outputVars, varName), response = "cardioDeath")
    curLM = lm(formula, dfCut)
    comparisonTab = anova(mainOnlyMod, curLM)
    
    pval = comparisonTab$`Pr(>F)`[2]
    if(pval<.05){
      aovTab = anova(curLM)
      ei = which(row.names(aovTab) == 'Residuals')
      ri = which(row.names(aovTab) == varName)
      etaVals[varName] = aovTab$`Sum Sq`[ri] /
              (aovTab$`Sum Sq`[ri] + aovTab$`Sum Sq`[ei])
    } else if(pval<.1){
      possibleVars = c(possibleVars, varName)
      print(comparisonTab)
    }
    
  }
  
  if(max(etaVals>0)){
    #get the best predictor and add it into the model
    ii = which(max(etaVals)==etaVals)
    bestVar = predNames[ii]
    predNames = predNames[-ii]
    outputVars = c(outputVars, bestVar)
    formula <- reformulate(outputVars, response = "cardioDeath")
    outputLM = lm(formula, dfCut)
  } else {
    check = F
  }
}

```

## Let's test for interactions with 3 good variables and other marginal ones

```{r interactions}

outputVars
possibleVars

#adding lostTeeth X noInternet interaction
exampInt <- lm(cardioDeath ~ lostTeeth + noVeg + smokeBan + lostTeeth*noInternet, 
               data = dfCut)

summary(exampInt)


#adding noVeg X noInternet interaction
exampInt <- lm(cardioDeath ~ lostTeeth + noVeg + smokeBan + noVeg*noInternet,  
               data = dfCut)

summary(exampInt)

#adding smokeBan X noInternet interaction
exampInt <- lm(cardioDeath ~ lostTeeth + noVeg + smokeBan + smokeBan*noInternet, 
               data = dfCut)

summary(exampInt)

```

## Plotting to see what the interaction might mean

```{r int plot}

sumDat = data.frame(cardioDeath = dfCut$cardioDeath, 
                    intModelResi = exampInt$residuals,
                    mainOnlyResi = mainOnlyMod$residuals, 
                    fitted = mainOnlyMod$fitted.values,
                    smokeBan = dfCut$smokeBan,
                    noInternet = dfCut$noInternet,
                    dummy = as.factor(1))

#let's cover all the space of possible smokeBan and noInternet values: 
test = expand.grid(seq(min(dfCut$smokeBan), 
                       max(dfCut$smokeBan), 
                       length.out = 50),
                   seq(min(dfCut$noInternet), 
                       max(dfCut$noInternet), 
                       length.out = 50))
names(test) = c('smokeBan', 'noInternet')
test = as.data.frame(test)
#for our fake data, let's hold lostTeeth and noVeg constant at their means
test$lostTeeth = mean(dfCut$lostTeeth)
test$noVeg = mean(dfCut$noVeg)
#we'll use the interaction model to predict outcomes by smokeBan and noInternet
test$cardioDeathPred = predict(exampInt, test)

ggplot(test, aes(x = smokeBan, y = noInternet, z = cardioDeathPred))+
  geom_contour_filled() + myTheme

#then let's plot our usual model plots: 

plotDat = sumDat %>% pivot_longer(cols = c('mainOnlyResi', 'intModelResi'), 
                                  names_to = 'model', 
                                  values_to = 'resVals')

plotDat %>% ggplot(aes(x = cardioDeath, y = resVals, color = model)) + 
  geom_point() + 
  geom_abline(slope = 0, intercept = 0,  
              color = 'red', linetype = "dashed", size = 1.5)+ 
  brightCol + myTheme 

plotDat %>% ggplot(aes(x = resVals, color = model, fill = model)) + 
  geom_histogram() + brightFil + 
  brightCol + myTheme + 
  facet_grid(model~.)



```

## Can we systematically test for interactions and add them to the model?

```{r systematic int}

workingMod = mainOnlyMod
check = T
interactionPairs <- combn(c(outputVars,possibleVars), 2, simplify = FALSE)
intNames = rep('a', length(interactionPairs))
for( ii in 1:length(interactionPairs)){
  pair = interactionPairs[[ii]]
  intNames[ii] <- paste0(pair[1], ":", pair[2])
}
while(check){

  etaVals = rep(0, length(intNames))
  names(etaVals) = intNames

for (interaction_term in intNames) {
  
  full_formula <- reformulate(c(outputVars, interaction_term), 
                              response = "cardioDeath")
  interaction_model <- lm(full_formula, dfCut)
  
  comparisonTab <- anova(workingMod, interaction_model)
  cat("\nTesting interaction:", interaction_term, "\n")
  
  pval = comparisonTab$`Pr(>F)`[2]
  if(pval<.05){
    # print(comparisonTab)
    aovTab = anova(interaction_model)
    ei = which(row.names(aovTab) == 'Residuals')
    ri = which(row.names(aovTab) == interaction_term)
    etaVals[interaction_term] = aovTab$`Sum Sq`[ri] /
            (aovTab$`Sum Sq`[ri] + aovTab$`Sum Sq`[ei])
  }
  
}

 if(max(etaVals>0)){
    #get the best predictor and add it into the model
    ii = which(max(etaVals)==etaVals)
    bestVar = intNames[ii]
    intNames = intNames[-ii]
    outputVars = c(outputVars, bestVar)
    formula <- reformulate(outputVars, response = "cardioDeath")
    workingMod = lm(formula, dfCut)
  } else {
    check = F
  }
}

```

## Interrogating our final interactive model, do we gain explanation power?

```{r plotting int mod}

summary(workingMod)

#let's cover all the space of possible lostTeeth and noVeg values: 
test = expand.grid(seq(min(dfCut$lostTeeth), 
                       max(dfCut$lostTeeth), 
                       length.out = 50),
                   seq(min(dfCut$noVeg), 
                       max(dfCut$noVeg), 
                       length.out = 50))
names(test) = c('lostTeeth', 'noVeg')
test = as.data.frame(test)
#for our fake data, let's hold lostTeeth and noVeg constant at their means
test$smokeBan = mean(dfCut$smokeBan)
test$mentalDistress = mean(dfCut$mentalDistress)
test$depression = mean(dfCut$depression)
#we'll use the interaction model to predict outcomes by smokeBan and noInternet
test$cardioDeathPred = predict(workingMod, test)

ggplot(test, aes(x = lostTeeth, y = noVeg, z = cardioDeathPred))+
  geom_contour_filled() + myTheme

ggplot(dfCut, aes(x = lostTeeth, y = noVeg))+
  geom_point(color = 'green') + myTheme

#how much variability did this interaction explain? partial Eta Squared
aovTab = Anova(workingMod, type = 3)
ei = which(row.names(aovTab) == 'Residuals')
ri = which(row.names(aovTab) == 'lostTeeth:noVeg')
etaVal = aovTab$`Sum Sq`[ri] /
      (aovTab$`Sum Sq`[ri] + aovTab$`Sum Sq`[ei])
print(paste0('The interaction between lostTeeth and noVeg had ',
             'partial eta squared of: ', round(etaVal, 3)))

#let's cover all the space of possible mentalDistress and depression values: 
test = expand.grid(seq(min(dfCut$mentalDistress), 
                       max(dfCut$mentalDistress), 
                       length.out = 50),
                   seq(min(dfCut$depression), 
                       max(dfCut$depression), 
                       length.out = 50))
names(test) = c('mentalDistress', 'depression')
test = as.data.frame(test)
#for our fake data, let's hold lostTeeth and noVeg constant at their means
test$smokeBan = mean(dfCut$smokeBan)
test$lostTeeth = mean(dfCut$lostTeeth)
test$noVeg = mean(dfCut$noVeg)
#we'll use the interaction model to predict outcomes by smokeBan and noInternet
test$cardioDeathPred = predict(workingMod, test)

ggplot(test, aes(x = mentalDistress, y = depression, z = cardioDeathPred))+
  geom_contour_filled() + myTheme

ggplot(dfCut, aes(x = mentalDistress, y = depression))+
  geom_point(color = 'green') + myTheme

#how much variability did this interaction explain? partial Eta Squared
aovTab = Anova(workingMod, type = 3)
ei = which(row.names(aovTab) == 'Residuals')
ri = which(row.names(aovTab) == 'mentalDistress:depression')
etaVal = aovTab$`Sum Sq`[ri] /
      (aovTab$`Sum Sq`[ri] + aovTab$`Sum Sq`[ei])
print(paste0('The interaction between mentalDistress and depression had ',
             'partial eta squared of: ', round(etaVal, 3)))

```

## Let's see how the addition of each term helped the overall model

```{r R2 v parameters}

RsqAdjVals <- numeric(length(outputVars))
modelTerms <- c()

for (ii in 1:length(outputVars)) {
  # Add the next variable
  modelTerms <- c(modelTerms, outputVars[ii])
  
  # Create formula string and model
  formula <- reformulate(modelTerms, response = "cardioDeath")
  model <- lm(formula, data = dfCut)
  
  # Extract adjusted R-squared
  RsqAdjVals[ii] <- summary(model)$adj.r.squared
}

# Make a data frame for plotting
rsq_df <- data.frame(
  num_terms = 1:length(outputVars),
  adj_r_squared = RsqAdjVals,
  dummy = as.factor(1)
)

# Plot

ggplot(rsq_df, aes(x = num_terms, y = adj_r_squared, color = dummy)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = 1:length(outputVars)) +
  labs(x = "Number of Parameters in Model",
       y = "Adjusted R²",
       title = "Adjusted R² vs. Model Complexity") +
  scale_y_continuous(breaks = seq(0,1,.1), limits = c(0,1)) +
  myTheme + brightCol + theme(legend.position = 'none')


```

