---
title: 'Combining Data from multiple Sources'
author: "Adam Dede"
date: "2024-01-02"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---

#### Examining the relationship between waste management and GDP

Data can be found here: 
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-05-21


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
myGitHubPath = 'C:\\Users\\Adam\\Documents\\GitHub\\MasterStatsUsingR\\'
# Load packages
# library(plyr)      # data visualization
library(tidyverse) # data manipulation and visualization
# library(rstatix)   # generating pipe-friendly functions for statistical analyses
# library(ggpubr)    # creating easily publication ready plots
library(janitor) #helps to rename titles
library(tidytuesdayR) #package for loading tidy tuesday datasets 
```

## Load Data in three CSVs and use janitor to make names R appropriate

```{r load data}

# Load the data 
#coastal vs mismanaged waste
cw <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/coastal-population-vs-mismanaged-plastic.csv")

#mismanaged waste vs gdp
mg <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/per-capita-mismanaged-plastic-waste-vs-gdp-per-capita.csv")

#total waste vs. gdp
wg <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/per-capita-plastic-waste-vs-gdp-per-capita.csv")

# inspect the data 
head(cw)
head(mg)
head(wg)
names(cw)
cw <- janitor::clean_names(cw)
names(cw)
names(mg)
mg <- janitor::clean_names(mg)
names(mg)
names(wg)
wg <- janitor::clean_names(wg)
names(wg)



```

## explore where valid data are

```{r examine valid data distribution}

# summarize how many good overvations there are for each variable of interest

#key variable in coast_vs_waste is the mismanaged plastic
dataCount_cw <- cw %>%
  select(year, mismanaged_plastic_waste_tonnes) %>%
  group_by(year) %>%
  summarize(naCount = sum(is.na(mismanaged_plastic_waste_tonnes)),
            n = n()) %>% 
  mutate(goodVals = n - naCount)

#key variable in mismanaged_vs_gdp is per captia mismanaged plastic
dataCount_mg <- mg %>%
  select(year, 
         per_capita_mismanaged_plastic_waste_kilograms_per_person_per_day) %>%
  group_by(year) %>%
  summarize(naCount = sum(is.na(
    per_capita_mismanaged_plastic_waste_kilograms_per_person_per_day)),
    n = n()) %>% 
  mutate(goodVals = n - naCount)

#key variable in waste_vs_gdp is total per capital plastic waste
dataCount_wg <- wg %>%
  select(year, 
         per_capita_plastic_waste_kilograms_per_person_per_day) %>%
  group_by(year) %>%
  summarize(naCount = sum(
    is.na(per_capita_plastic_waste_kilograms_per_person_per_day)), 
    n = n()) %>% 
  mutate(goodVals = n - naCount)


dataCount_cw %>% ggplot(aes(x = year-.5, y = goodVals)) +
  geom_point(color = 'green') + 
  geom_point(data = dataCount_mg, aes(x = year, y = goodVals), color = 'red')+
  geom_point(data = dataCount_wg, aes(x = year+.5, y = goodVals), 
             color = 'black') +
  xlim(c(2000, 2015))


```

## remove missing data rows, select target year, and merge into a single data frame

```{r get rid of NA and select year}
#down select to remove missing and select only 2010
cw <- cw %>% filter(!is.na(mismanaged_plastic_waste_tonnes), year == 2010)
mg <- mg %>% filter(!is.na(
        per_capita_mismanaged_plastic_waste_kilograms_per_person_per_day),
        year == 2010)
wg <- wg %>% filter(!is.na(
        per_capita_plastic_waste_kilograms_per_person_per_day), 
        year ==2010)



df <- merge(cw, mg, by = "code", suffixes = c("_df1", "_df2"))
df <- df %>%
  select(-ends_with("_df2"))
new_column_names <- gsub("_df1$", "", names(df))
names(df) <- new_column_names

df <- merge(df, wg, by = "code", suffixes = c("_df1", "_df2"))
df <- df %>%
  select(-ends_with("_df2"))
new_column_names <- gsub("_df1$", "", names(df))
names(df) <- new_column_names
```

## rename variables to make them easier to work with

```{r make easier names}

names(df)
colnames(df) <- c('country', 'countryLong', 'year', 'mismanagedTotal',
                  'coastalPop', 'totalPop', 'mismanagedPerCap', 'gdpPerCap',
                  'totWastePerCap', 'gdpPerCap2')
names(df)
df <- df[,-10]

df <- na.omit(df)
  
```

# Data Visualisation: Histograms 

Create histograms to display the distributions of all variables in the data

```{r histograms}

df %>% ggplot(aes(x = mismanagedTotal))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = coastalPop))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = totalPop))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = mismanagedPerCap))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = gdpPerCap))+
  geom_histogram() +
theme_classic()

df %>% ggplot(aes(x = totWastePerCap))+
  geom_histogram() +
theme_classic()


```
## save our cleaned data frame for later use! 
```{r save the clean .csv}
write.csv(df, paste0(myGitHubPath, 'data\\plasticWaste.csv'))

```
