---
title: 'Arrays and Matrices'
Section: 'Basic R Coding'
Author: "Adam Dede"
Course: "Data to Decisions: Master Statistics with R and tidyverse"
Course_URL: "https://www.udemy.com/course/statsr_ad" 
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)



```

## creating and indexing in matrices works like vectors but with two dimensions

```{r make a matrix}
x = matrix(c(1,2,3,4), nrow = 2, ncol = 2)

x[1,]
x[,1]

y = matrix(nrow = 2, ncol = 2)
 y[1,2] = 1

```
## switching to row major orientation
```{r populating matrices along rows}
z = matrix(c(1,2,3,4,5,6), nrow = 2, byrow = T)
z 


```

## you can do math operations on whole matrices like on vectors with some extras
```{r matrix math}
x + x

x*3

x %*% z



```

## matrices are vectors with row and column attributes
```{r matrix vector relationship}
attributes(z)
length(z)

z[4]

z = c(z)
attributes(z)
z
z[4]


```

## adding rows and columns to matrices
```{r rows and columns}
cbind(x, c(100, 100))
rbind(x, c(100, 100))

cbind(x, c('hello', 'there'))
rbind(x, c(T, F))



```
## recycling in matrices
```{r recycling}
matrix(c(1,2), nrow = 3, ncol = 2)
matrix(c(1,2), nrow = 3, ncol = 2, byrow = T)

#can you use recycling to make a matrix with nrow = 5, ncol = 5
#each col is 1,2,3,4,5? 

# your code here 


```
##dimensions can get dropped and break your code
```{r dimension dropping}
r = x[1,]
str(r)
str(x)

r = x[1,,drop = F]



```

## arrays (matrices with more than 2 dimensions)
```{r arrays}
a = array(c(1:18), dim = c(2,3,3))


```
