---
title: 'Arrays and Matrices'
Section: 'Basic R Coding'
Author: "Adam Dede"
Course: "Data to Decisions: Master Statistics with R and tidyverse"
Course_URL: "..." 
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---


## set environment

```{r message = FALSE, warning = FALSE}

# Set rendering parameters
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)



```

## creating and indexing in matrices works like vectors but with two dimensions

```{r make a matrix}


```
## switching to row major orientation
```{r populating matrices along rows}




```

## you can do math operations on whole matrices like on vectors with some extras
```{r matrix math}




```

## matrices are vectors with row and column attributes
```{r matrix vector relationship}


```

## adding rows and columns to matrices
```{r rows and columns}


```
## recycling in matrices
```{r recycling}

```
##dimensions can get dropped and break your code
```{r dimension dropping}

```

## arrays (matrices with more than 2 dimensions)
```{r arrays}



```
